---
title: "ParallelCordinates"
author: "TracyLiu"
date: "2020/11/18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(GGally)
library("parcoords")
library("d3r")
```

# Analyzing with 5 big occupation catogories
```{r,fig.width=6,fig.height=8}
YearBoroughData <- read.csv("/Users/tracy/Desktop/columbia/5702/STAT5702_NYC_Employment_Analysis/DataPreprocessing/Data/Preprocess_ClevelandDotPlot/OccupationByYearBorough.csv")
YearBoroughData$X <- NULL

YearBoroughData.tidy <- pivot_longer(YearBoroughData,cols = !c("Occupations","year"), names_to = "boroughs", values_to = "NumEmp")
YearBoroughData.wide <- pivot_wider(YearBoroughData.tidy, names_from = Occupations, values_from = NumEmp)

YearBoroughData.wide$year <- NULL

parcoords(YearBoroughData.wide,
          rownames = F,
          brushMode = "1D-axes",
          reorderable = T,
          queue = T,
          color = list(
            colorBy = "boroughs",
            colorScale = "scaleOrdinal",
            colorScheme = "schemeCategory10"
          ),
          withD3 = TRUE) 
```
From the plot above, we discover that
1) Management, business, science, and arts occupations and Service occupations: for Bronx, Queens and Manhattan, the two types of occupations are positively related with each other, while for Kings, the two types of occupations are negatively corelated.
2) Service occupations and Sales and office occupations: for Manhattan and Richmond, the two types of occupations are negatively corelated. For other boroughs, there are not significant relations.
3) Sales and office occupations and Natural resources, construction, and maintenance occupations: all boroughs show a negative corelation.
4) Natural resources, construction, and maintenance occupations and Production, transportation, and material moving occupations: all boroughs show a negative corelation.

# Analyzing with 25 small occupation catogories
  * There might be too many categories
```{r,fig.width=6,fig.height=8}
YearBoroughData <- read.csv("/Users/tracy/Desktop/columbia/5702/STAT5702_NYC_Employment_Analysis/DataPreprocessing/Data/Preprocess_ClevelandDotPlot/OccupationByYearBorough_small.csv")
YearBoroughData$X <- NULL

YearBoroughData.tidy <- pivot_longer(YearBoroughData,cols = !c("Occupations","year"), names_to = "boroughs", values_to = "NumEmp")
YearBoroughData.wide <- pivot_wider(YearBoroughData.tidy, names_from = Occupations, values_from = NumEmp)

YearBoroughData.wide$year <- NULL

parcoords(YearBoroughData.wide,
          rownames = F,
          brushMode = "1D-axes",
          reorderable = T,
          queue = T,
          color = list(
            colorBy = "boroughs",
            colorScale = "scaleOrdinal",
            colorScheme = "schemeCategory10"
          ),
          withD3 = TRUE) 
```