---
title: "EducationOverTime.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r imports}
library(RCurl)
```

```{r Data}
Education <- read.csv(text = getURL("https://raw.githubusercontent.com/tracy3057/STAT5702_NYC_Employment_Analysis/main/DataPreprocessing/Data/education/Education.csv"))
PropBachelor <- read.csv(text = getURL("https://raw.githubusercontent.com/tracy3057/STAT5702_NYC_Employment_Analysis/main/DataPreprocessing/Data/education/PropBachelor.csv"))
PropBachelorByGender <- read.csv(text = getURL("https://raw.githubusercontent.com/tracy3057/STAT5702_NYC_Employment_Analysis/main/DataPreprocessing/Data/education/PropBachelorByGender.csv"))
```

```{r Proportion College Educated Over Time}
ggplot(data = PropBachelor,aes(Year,PropBachelor)) + geom_line(size = 1, col = 'black') + xlab('Year') + ylab('Proportion With Bachelors Degree') + scale_x_continuous(breaks = seq(2010,2019)) + geom_point(size = 2.5, col = 'royalblue3') + ylim(0,.5)
```
Observation of Bachelor's proportions over time:

It appears that the proportion of people with Bachelor's Degrees is steadily increasing over time. In the last 10 years the rate has increased from 30% to 35% which indicates an average increase of about half a percent per year.

```{r Proportion College Educated Over Time by Gender}
PropBachelorByGenderTidy <- pivot_longer(PropBachelorByGender[,1:3],!Year,names_to = 'Gender',values_to = 'Proportion')

ggplot(PropBachelorByGenderTidy) + geom_line(aes(Year,Proportion,group = Gender, color = Gender), size = 1) + geom_point(aes(Year,Proportion,group = Gender, color = Gender), size = 5) + xlab('Year') + ylab('Proportion With Bachelors Degree') + scale_x_continuous(breaks = seq(2010,2019)) + ylim(0,.5)

ggplot(data = PropBachelorByGender,aes(Year,Difference)) + geom_line(size = 1, col = 'black') + xlab('Year') + ylab('Difference between Male and Female with Bachelors') + scale_x_continuous(breaks = seq(2010,2019)) + geom_point(size = 2.5, col = 'royalblue3') + ylim(0,.1)
```
Observations from Proportion of Males and Females with Bachelors Degrees from 2010-2019:

It seems that in 2010, a larger proportion of females than men had bachelor's degrees. While the proportions are increasing over time for both males and females, it seems that the rate at which a higher proportion of females have bachelors degree is actually faster than it is for males. 

This is confirmed by the gap between the male and female dots on the graph continuing to grow over the course of this decade, and the next plot confirms this.

Observations from Difference in Proportion of Females and Males with a Bachelor's Degree from 2010-2019:

As was able to be seen from the previous plot, the gap between female and male education proportion is growing over time. If trends continue we may see a future where females are significantly more educated than males, even more than they currently are.
