## GDP

```{r,include=FALSE,echo=FALSE}
library(tidyverse)
library(RCurl)
library(scales)
Salary_NumEmp <- read.csv(text = getURL("https://raw.githubusercontent.com/tracy3057/STAT5702_NYC_Employment_Analysis/main/DataPreprocessing/Data/gdp/NYCSalary_gdp.csv"))
Salary_NumEmp$X <- NULL
```

```{r,include=FALSE,echo=FALSE}
# NumEmp
Salary_NumEmp <- Salary_NumEmp %>% select(1,2,5)
```

```{r,include=FALSE,echo=FALSE}
Salary_NumEmp[Salary_NumEmp$Earning_category=="FULL-TIME YEAR-ROUND WORKERS WITH EARNINGS","Earning_category"] <- "Full-time year-round workers with earnings"

Salary_NumEmp[Salary_NumEmp$Earning_category=="Mean earnings (dollars) for full-time year-round workers with earnings","Earning_category"] = "Mean earnings (dollars)"

Salary_NumEmp$Earning_category <- as.factor(Salary_NumEmp$Earning_category)

Salary_NumEmp <- Salary_NumEmp %>% pivot_wider(names_from = "Earning_category", values_from = "New.York..Total")
# add GDP values
Salary_NumEmp$NYCGDP <- c(65438,65092,67565,67227,68661,69814,70781,72434,73508,75131)
Salary_NumEmp$GDP <- c(48468,49887,51611,53118,55048,56823,57928,59958,62840,65118)

Salary_NumEmp <- Salary_NumEmp %>% pivot_longer(cols = c(5:13),names_to = "Earning_Category", values_to = "Value")

Salary_NumEmp$Earning_Category = factor(Salary_NumEmp$Earning_Category, levels=c('$1 to $9999 or loss','$10000 to $14999','$15000 to $24999','$25000 to $34999','$35000 to $49999','$50000 to $64999','$65000 to $74999','$75000 to $99999','$100000 or more'))
```

```{r scale,include=FALSE,echo=FALSE}
ks <- function (x) { number_format(accuracy = 1,
                                   scale = 1/1000,
                                   suffix = "k",
                                   big.mark = ",")(x) }
```

```{r,fig.width=14,fig.height=9,include=FALSE,echo=FALSE}
# plot
ggplot(Salary_NumEmp,aes(NYCGDP,Value))+
  geom_point(color = "royalblue3", size = 2, alpha = 0.75)+
  scale_x_continuous(labels = ks)+
  scale_y_continuous(labels = ks)+
  xlab("NYC GDP Per Capita")+
  ylab("NumEmp")+
  ggtitle("NumEmp vs NYC GDP Per Capita")+
    theme(axis.text=element_text(size=8),
      axis.title=element_text(size=15,face="bold"),
        strip.text.x = element_text(
        size = 12, face = "bold.italic"
        ),
      strip.text.y = element_text(
        size = 12, color = "red", face = "bold.italic"
        ),
      plot.title = element_text(size = 18, face = "bold"))+
  facet_wrap(~Earning_Category,scales = "free_y")
```

```{r,fig.width=14,fig.height=9}
ggplot(Salary_NumEmp,aes(GDP,Value))+
  geom_point(color = "royalblue3", size = 2, alpha = 0.75)+
  scale_x_continuous(labels = ks)+
  scale_y_continuous(labels = ks)+
  xlab("GDP Per Capita")+
  ylab("NumEmp")+
  ggtitle("NumEmp vs GDP Per Capita")+
      theme(axis.text=element_text(size=8),
      axis.title=element_text(size=15,face="bold"),
        strip.text.x = element_text(
        size = 12, face = "bold.italic"
        ),
      strip.text.y = element_text(
        size = 12, color = "red", face = "bold.italic"
        ),
      plot.title = element_text(size = 18, face = "bold"))+
  facet_wrap(~Earning_Category,scales = "free_y")
```

**Observations on NumEmp vs GDP/NYC GDP Per Capita:**

The scatter dot plots above show us the relationship between employment and GDP/NYC GDP per capita in different salary categories. From the plots, we can see clearly that there is a turning point in the categories of '\$25000 to \$34999','\$35000 to \$49999'. Before the two categories, the number of employed is in a decreasing trend when GDP/NYC GDP per capita goes up. However, in the groups with higher salaries, the number of employed is in a increasing trend when GDP/NYC GDP per capita goes up.


